### 2016-01-22 Manuscript/BiologyLetters/response.tex

\item{\textcolor{blue}{ESM 1, p. 3: ``one of these repository'' should be ``one of these repositories.''}}
We fixed this typo.

\item{\textcolor{blue}{ESM 1, p. 4: ``that were to irrelevant'' should be ``that were too irrelevant.''}}
We fixed this typo.

\item{\textcolor{blue}{ESM 1, p. 4: ``of the recent published matrix'' should be ``of the recently published matrices.''}}
We fixed this typo.

\item{\textcolor{blue}{ESM 1, p. 4: First sentence of the paragraph in the middle of the page is a sentence fragment, and needs to state where the matrices are available.}}
We added the two missing link in the ESM1: ``The list of all the 286 download matrices is available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices}.'' and ``All the standardised matrices are available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices_binomial/Matrices}.''

@@

\item{\textcolor{blue}{Reviewer 3 (Pete Wagner)’s major problem - with which I agree - is semantic, and thus easily fixable with a few explicit caveats (e.g., ``by cladistic we mean...'').}}

\item{\textcolor{blue}{One point I would add of my own is that the orders used are obviously very variable in size (species number), so what about total coverage?
This could surely be synthesised from the tables and would be a nice simple way to compare the with and without threshold results.
Similarly, I think Reviewer 3’s [We assume Graeme Lloyd is refering to Graham Slater's reviewe (reviewer 2)] %or same as comment before: no names!
 graphs are a nice potential addition to the ESM.}}
\end{enumerate}

\subsection{Reviewer 4 specific comments:}
\begin{enumerate}
\item{\textcolor{blue}{Lines 72-74 - Web addresses should be consistent in dropping of ``http://www.''.}}
We removed the ``http://www.'' part of the URLs on p.@, lines@ and in the ESM1.

\item{\textcolor{blue}{Line 104 - ``NTI 15'' should probably be ``NTI [15]''.}}
%TO DO IN THE WORD MANUSCRIPT! We fixed this typo.

\item{\textcolor{blue}{Line 104 - It’s not completely clear to me how ``distance'' is measured. Is this branch-lengths on a time-scaled tree? Or just a node count?}}
We changed ``$\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa'' to ``$\overline{MNND}_{obs}$ is the observed mean branch length between each of $n$ taxa'' and ``$\overline{MPD}_{obs}$ is the observed mean phylogenetic distance of the tree containing only the $n$ taxa'' to ``$\overline{MPD}_{obs}$ is the observed mean phylogenetic branch length of the tree containing only the $n$ taxa'' on p.@, lines@.

\item{\textcolor{blue}{Line 117 - It’s not clear to me how supraspecific values were calculated if the reference phylogeny (ref 10) is species-level. Are you using the node subtending the higher taxon or something else?}}
%TO CHECK!

\item{\textcolor{blue}{Line 149 - ``most palaeontological studies use the genus as their smallest OTU'' I’m not sure if this is strictly true. It’s more that shorter OTU names are commonly used. After all at least one actual specimen has to be coded and in my experience if you read the paper usually only a single species is used. Palaeontologists are just bad at using proper binomials in their matrices (which is an issue, but a different one). In addition ref 20 is actually all species-level OTUs! (I didn’t check ref 21.)}}
%To CHECK, do we actually care?

\item{\textcolor{blue}{Line 154 - ``phylogenetically overdispersed'' Purely semantic, but I guess what you mean is that even sampling would be ideal, and this is *measured* using phylogenetic over-dispersion.}}
We changed ``the ideal scenario is for them to be phylogenetically overdispersed'' to ``the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion)'' on p.@, line@.

\item{\textcolor{blue}{Line 245 - The caption could be more informative regarding the random and clustered distributions that these figures represent.}}
We changed the caption to: ``Phylogenetic distribution of species with available cladistic data across two orders (A: Primates; B: Carnivora). Blue branches indicate available cladistic data for the species. Cladistic data is randomly distributed in Primates (A) and clustered in Carnivora (B).''.

\item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``The list of all the 286 download matrices is available on''}}
See reviewer 1 specific comment 12.

\item{\textcolor{blue}{ESM 1 - Incomplete sentence: ``All the standardised matrices are available on''}}
See reviewer 1 specific comment 12.

\item{\textcolor{blue}{ESM 1 - ``designated as ``fossil'' all OTUs that were present in the Paleobiology database'' The database contains extant taxa too, although these should be marked as such. I guess this doesn’t matter given the flow chart in Figure 2, but is something the authors should be aware of. (Obviously more generally living species can have fossil records, or at least plenty of fossils are assigned to extant species.)}}
%TO DO: change ``fossil'' to ``non living OTU''? As in OTU that is not present in the Wilson Reeder list or the Bininda Emonds tree?

\item{\textcolor{blue}{ESM 1 Figure 2 - How many OTUs were ignored? And based on the (assumed) cursory inspection of the authors what are they? Higher taxa? Specimens? Or just species you can’t reconcile with your three data sets?}}
%TO DO: add the number of ignored OTUs. I think they're mainly specimen IDs without species names.

\item{\textcolor{blue}{ESM 2 - ``The following section contains supplementary results to the main body: the available data structure using the NTI and the PD metric'' Abbreviations should be introduced when first used, not later as they are here.}}
We changed ``using the NTI and the PD metric'' to ``using the Net Relatedness Index (NRI) and the Nearest Taxon Index (NTI) metrics''.
\end{enumerate}

### 2016-01-22 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Minor comments (typos) fixed

%---------------------------------------------
%
%       METHODS
%
%---------------------------------------------
\section{Materials and Methods}
\subsection{Data collection and standardisation}
We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{http://www.morphobank.org/} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{https://github.com/rossmounce/cladistic-data}).
We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for a detailed description of the search procedure).
In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
We used OTUs rather than species because% rev 1.2.3
 entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.

@@

NTI \cite{webb2002phylogenies} is based on mean nearest neighbour distance ($MNND$) and is calculated as follows:
  \begin{equation}
    \text{NTI}=-\left(\frac{\overline{MNND}_{obs}-\overline{MNND}_{n}}{\sigma(MNND_{n})}\right)
  \end{equation}
where $\overline{MNND}_{obs}$ is the observed mean distance between each of $n$ taxa with cladistic data and its nearest neighbour with cladistic data in the phylogeny, 
$\overline{MNND}_{n}$ is the mean of 1000 $MNND$ between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random $MNND$ values.
NRI is calculated in the same way, but using the mean phylogenetic distance ($MPD$):
  \begin{equation}
    \text{NRI}=-\left(\frac{\overline{MPD}_{obs}-\overline{MPD}_{n}}{\sigma(MPD_{n})}\right)
  \end{equation}
where $\overline{MPD}_{obs}$ is the observed mean phylogenetic distance of the tree containing only the $n$ taxa with cladistic data.
Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values reflect clustering.

@@

When few species have available cladistic data, the ideal scenario is for them to be phylogenetically overdispersed to maximize the possibilities of a fossil branching in the right clade.
The second best scenario is that species with cladistic data are randomly distributed across the phylogenies. 
Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny.
The worst case scenario for fossil placement is that species with cladistic data are phylogenetically clustered. 
Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance. 
Our results suggest that this may be problematic at the genus-level in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade% rev 1.2.7
, and will have more chance to randomly branch within Canidae (Figure \ref{Figure_example_coverage}B).

>>>>

%---------------------------------------------
%
%       METHODS
%
%---------------------------------------------
\section{Materials and Methods}
\subsection{Data collection and standardisation}
We downloaded all cladistic matrices containing any living and/or fossil mammal taxa from three major public databases: MorphoBank (\url{morphobank.org} \cite{morphobank}), Graeme Lloyd's website (\url{graemetlloyd.com/matrmamm.html}) and Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}).
We also performed a systematic Google Scholar search for matrices that were not uploaded to these databases (see Electronic Supplementary Material (ESM) for a detailed description of the search procedure).
In total, we downloaded 286 matrices containing 5228 unique operational taxonomic units (OTUs). 
We used OTUs rather than species because% rev 1.2.3
 entries in the matrices ranged from species to families, and standardised the taxonomy as described in the ESM. 
We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{BinindaEmonds} or the taxonomy of \cite{wilson2005mammal}.

@@

NTI \cite{webb2002phylogenies} is based on mean nearest neighbour distance ($MNND$) and is calculated as follows:
  \begin{equation}
    \text{NTI}=-\left(\frac{\overline{MNND}_{obs}-\overline{MNND}_{n}}{\sigma(MNND_{n})}\right)
  \end{equation}
where $\overline{MNND}_{obs}$ is the observed mean branch length between each of $n$ taxa with cladistic data and its nearest neighbour with cladistic data in the phylogeny, 
$\overline{MNND}_{n}$ is the mean of 1000 $MNND$ between $n$ randomly drawn taxa, and $\sigma(MNND_{n})$ is the standard deviation of these 1000 random $MNND$ values.
NRI is calculated in the same way, but using the mean phylogenetic distance ($MPD$):
  \begin{equation}
    \text{NRI}=-\left(\frac{\overline{MPD}_{obs}-\overline{MPD}_{n}}{\sigma(MPD_{n})}\right)
  \end{equation}
where $\overline{MPD}_{obs}$ is the observed mean phylogenetic branch length of the tree containing only the $n$ taxa with cladistic data.
Negative NTI and NRI values show that the focal taxa are more overdispersed across the phylogeny than expected by chance, and positive values reflect clustering.

@@

When few species have available cladistic data, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil branching in the right clade.
The second best scenario is that species with cladistic data are randomly distributed across the phylogenies. 
Here we expect no special bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with cladistic data were randomly distributed across the phylogeny.
The worst case scenario for fossil placement is that species with cladistic data are phylogenetically clustered. 
Then we expect two major biases to occur: first, fossils will not be able to branch within a clade containing no data, and second, fossils will have higher probability of branching within the most sampled clade by chance. 
Our results suggest that this may be problematic at the genus-level in Carnivora, Cetartiodactyla, Chiroptera and Soricomorpha. 
For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade% rev 1.2.7
, and will have more chance to randomly branch within Canidae (Figure \ref{Figure_example_coverage}B).

### 2016-01-22 Manuscript/BiologyLetters/Supplementary/ESM_1.tex
Minor comments (typos) fixed

\subsubsection{Morphobank}
We accessed the Morphobank repository (\texttt{http://www.morphobank.org/}) on the 10th of June 2015 and used the keywords listed above in the search menu.
We downloaded the data associated with each project matching with the keyword.

\subsubsection{Graeme Lloyd}
We accessed Graeme Lloyd's website repository (\texttt{http://graemetlloyd.com/}) on the 10th of June 2015 and downloaded all the matrices that were available with a direct download link in the mammal data section of the website (\texttt{http://graemetlloyd.com/matrmamm.html}).

\subsubsection{Ross Mounce}
We accessed Ross Mounce's GitHub repository (\texttt{https://github.com/rossmounce}) on the 11th of June 2015 and downloaded all 601 matrix.
We then ran a shell script to select only the matrices that had any text element that match with one of the search terms.
% LINK WHICH SHELL SCRIPT
To make the matrix selection more thorough, we ignored the keywords case as well as the latin suffix (\textit{ia}, \textit{ata}, \textit{ea}, and \textit{a}).

@@

The list of all the 286 download matrices is available on \url{https://github.com/TGuillerme/Missing_living_mammals}.
The matrices contained a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique.
In this study, we refer to OTUs rather than species because the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e. the name of a collection item) to the family-level.
Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
Therefore for some orders, we sampled more genera than species.

\subsection{Standardising the matrices}
We transformed all the non-nexus matrices (tnt, word, excel, jpeg) to nexus format manually.
We then cleaned the nexus matrices by removing any extra information (trees, continuous characters, morphological characters description, molecular data) to end up with nexus matrices containing only the discrete morphological data.
We then manually fixed the incorrectly-formatted bionomial names format (e.g. \textit{H. sapiens} became \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
All the standardised matrices are available on \url{https://github.com/TGuillerme/Missing_living_mammals}.

\subsection{Selecting the living OTUs}

To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H. sapiens} was transformed to \textit{Homo sapiens}).
We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{bininda-emondsthe2007} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
For OTUs that did not appear in these three sources, we first decomposed the name (i.e. \textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; Figure ~\ref{Supp_figure_Taxonomic_algorithm}).

@@

\section{2 - Data reproducibility}
Every step of the analysis (apart from downloading and standardisation of the matrices) is entirely repeatable via GitHub (\url{https://github.com/TGuillerme/Missing_living_mammals}).
%Needs to fix this in GitHub
%1-Rarefaction ?
%2-Selecting the OTUs
%3-Analysing the availability and the data structure

>>>>

\subsubsection{Morphobank}
We accessed the Morphobank repository (\url{morphobank.org}) on the 10th of June 2015 and used the keywords listed above in the search menu.
We downloaded the data associated with each project matching with the keyword.

\subsubsection{Graeme Lloyd}
We accessed Graeme Lloyd's website repository (\url{graemetlloyd.com/matrmamm.html}) on the 10th of June 2015 and downloaded all the matrices that were available with a direct download link in the mammal data section of the website (\texttt{http://graemetlloyd.com/matrmamm.html}).

\subsubsection{Ross Mounce}
We accessed Ross Mounce's GitHub repository (\url{github.com/rossmounce/cladistic-data}) on the 11th of June 2015 and downloaded all 601 matrix.
We then ran a shell script to select only the matrices that had any text element that match with one of the search terms.
% LINK WHICH SHELL SCRIPT
To make the matrix selection more thorough, we ignored the keywords case as well as the latin suffix (\textit{ia}, \textit{ata}, \textit{ea}, and \textit{a}).

@@

The list of all the 286 download matrices is available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices}.
The matrices contained a total of 11010 operational taxonomic units (OTUs) of which 5228 were unique.
In this study, we refer to OTUs rather than species because the entries in the downloaded matrices were not standardised and ranged from specific individual specimen names (i.e. the name of a collection item) to the family-level.
Where possible, we considered OTUs at their lowest valid taxonomic level (i.e. species) but some OTUs were only valid at a higher taxonomic level (e.g. genus or family).
Therefore for some orders, we sampled more genera than species.

\subsection{Standardising the matrices}
We transformed all the non-nexus matrices (tnt, word, excel, jpeg) to nexus format manually.
We then cleaned the nexus matrices by removing any extra information (trees, continuous characters, morphological characters description, molecular data) to end up with nexus matrices containing only the discrete morphological data.
We then manually fixed the incorrectly-formatted bionomial names format (e.g. \textit{H. sapiens} became \textit{Homo sapiens}) using the abbreviation list in the concerned publications. 
All the standardised matrices are available on \url{github.com/TGuillerme/Missing_living_mammals/tree/master/Data/Matrices_binomial/Matrices}.

\subsection{Selecting the living OTUs}
To select the lowest valid taxonomic level for each OTU, we standardised their taxonomy by correcting species names so they matched standard taxonomic nomenclature (e.g., \textit{H. sapiens} was transformed to \textit{Homo sapiens}).
We designated as ``living'' all OTUs that were either present in the phylogeny of \cite{bininda-emondsthe2007} or the taxonomy of \cite{wilson2005mammal}, and designated as ``fossil'' all OTUs that were present in the Paleobiology database (\url{https://paleobiodb.org/}).
For OTUs that did not appear in these three sources, we first decomposed the name (i.e. \textit{Homo sapiens} became \textit{Homo} and \textit{sapiens}) and tried to match the first element with a higher taxonomic level (family, genus etc.).
Any OTUs that still had no matches in the sources above were designated as non-applicable (NA; Figure ~\ref{Supp_figure_Taxonomic_algorithm}).

@@

\section{2 - Data reproducibility}
Every step of the analysis (apart from downloading and standardisation of the matrices) is entirely repeatable via GitHub (\url{github.com/TGuillerme/Missing_living_mammals}).
%Needs to fix this in GitHub
%1-Rarefaction ?
%2-Selecting the OTUs
%3-Analysing the availability and the data structure
